---
title: "Select New Experiments"
author: "Tugce Ozturk"
date: "January 21, 2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Load in libraries
```{r}
library(dplyr)
library(ggplot2)
library(factoextra)
library(NbClust)
library(GGally)
library(plotly)
library(corrplot)
```

## Load in Pareto table
```{r}
pareto.table <- readr::read_csv("paretoTable.csv", col_names = TRUE)
```


## Partition the dataset.  
```{r}
pareto.IO <- pareto.table%>% select(Ax, Bx, Dx,
         Cx, Fx, Ex, Gx, Hx, Ix, Jx, Kx, Lx, Mx, Nx)
colnames(pareto.IO) <- c("Ax", "Bx", "Dx", "Cx", "Fx", "Ex", "Gx","Hx", "Ix", "Jx","Kx","Lx", "Mx", "Nx")
```


## Distributions for each of the input and mean of variables. 
```{r}
ggplot(data = reshape2::melt(pareto.IO)) +
    geom_histogram(bins = 25, fill = "blue", mapping = aes(x = value)) +
    facet_wrap(~variable, scales = "free")+theme_bw()
```


## Standardize the variables
```{r}
df <- as.data.frame(scale(pareto.IO), center = TRUE, scale = TRUE)
df_output_mean <- df%>%select("Gx","Hx", "Ix", "Jx")
df_output_sd <- df%>%select("Kx","Lx", "Mx", "Nx")
df_output <- df%>%select("Gx","Hx", "Ix", "Jx","Kx","Lx", "Mx", "Nx")
```

## CLUSTERING BASED ON OUTPUT MEAN

## Select optimum number of clusters
```{r}
fviz_nbclust(NbClust(df_output_mean, distance = "euclidean", index = "all", method = "ward.D2"))
```

## Hierarchical clustering
```{r}
d <- dist(df_output_mean, method = "euclidean")
link.ward.D2 <- hclust(d, method = "ward.D2")
```

## Input boxplot by cluster
```{r}
clust.3 <- cutree(link.ward.D2, k = 3)
ggplot(data = reshape2::melt(pareto.IO %>% select("Ax", "Bx", "Dx", "Cx", "Fx", "Ex") %>% mutate(row.ID = 1:nrow(pareto.IO)),
                             id.vars = c("row.ID")) %>% 
           left_join(.,
                     data.frame(row.ID = 1:nrow(df),
                                cluster = clust.3),
                     by = c("row.ID"))) + 
    geom_boxplot( mapping = aes(x = cluster, y=value, fill = as.factor(cluster))) + 
    facet_wrap(~variable, scales = "free") + 
    scale_fill_discrete("Cluster")+theme_bw()
```

## Output boxplot by cluster
```{r}
clust.3 <- cutree(link.ward.D2, k = 3)
ggplot(data = reshape2::melt(pareto.IO %>% select("Gx","Hx", "Ix", "Jx") %>% mutate(row.ID = 1:nrow(pareto.IO)),
                             id.vars = c("row.ID")) %>% 
           left_join(.,
                     data.frame(row.ID = 1:nrow(df),
                                cluster = clust.3),
                     by = c("row.ID"))) + 
    geom_boxplot( mapping = aes(x = cluster, y=value, fill = as.factor(cluster))) + 
    facet_wrap(~variable, scales = "free") + 
    scale_fill_discrete("Cluster")+theme_bw()
```

## Matrix of scatterplots for visual inspection
```{r}
matrix.df <-df_output_mean
matrix.df$cluster <- as.factor(clust.3)
ggpairs(matrix.df,columns = 1:4, bins=10, aes(colour = cluster))
```



## Calculate the correlation matrix for output.  
```{r}
cor.matrix <- cor(df_output_mean)
corrplot::corrplot(cor.matrix, order = "hclust", addrect = 3)
```

## Calculate the correlation matrix for input+output.  
```{r}
cor.matrix <- cor(df)
corrplot::corrplot(cor.matrix, order = "hclust", addrect = 3)
```

## Hierarchical clustering dendogram
```{r}
fviz_dend(link.ward.D2, k = 3, rect = TRUE)
```


## CLUSTERING BASED ON OUTPUT STDEV

## Select optimum number of clusters
```{r}
fviz_nbclust(NbClust(df_output_sd, distance = "euclidean", index = "all", method = "ward.D2"))
```

## Hierarchical clustering
```{r}
d <- dist(df_output_sd, method = "euclidean")
link.ward.D2 <- hclust(d, method = "ward.D2")
```

## Input boxplot by cluster
```{r}
clust.4 <- cutree(link.ward.D2, k = 4)
ggplot(data = reshape2::melt(pareto.IO %>% select("Ax", "Bx", "Dx", "Cx", "Fx", "Ex") %>% mutate(row.ID = 1:nrow(pareto.IO)),
                             id.vars = c("row.ID")) %>% 
           left_join(.,
                     data.frame(row.ID = 1:nrow(df),
                                cluster = clust.4),
                     by = c("row.ID"))) + 
    geom_boxplot( mapping = aes(x = cluster, y=value, fill = as.factor(cluster))) + 
    facet_wrap(~variable, scales = "free") + 
    scale_fill_discrete("Cluster")+theme_bw()
```

## Output boxplot by cluster
```{r}
clust.4 <- cutree(link.ward.D2, k = 4)
ggplot(data = reshape2::melt(pareto.IO %>% select("Kx","Lx", "Mx", "Nx") %>% mutate(row.ID = 1:nrow(pareto.IO)),
                             id.vars = c("row.ID")) %>% 
           left_join(.,
                     data.frame(row.ID = 1:nrow(df),
                                cluster = clust.4),
                     by = c("row.ID"))) + 
    geom_boxplot( mapping = aes(x = cluster, y=value, fill = as.factor(cluster))) + 
    facet_wrap(~variable, scales = "free") + 
    scale_fill_discrete("Cluster")+theme_bw()
```

## Matrix of scatterplots for visual inspection
```{r}
matrix.df <-df_output_sd
matrix.df$cluster <- as.factor(clust.4)
ggpairs(matrix.df,columns = 1:4, bins=10, aes(colour = cluster))
```


## Calculate the correlation matrix for output.  
```{r}
cor.matrix <- cor(df_output_sd)
corrplot::corrplot(cor.matrix, order = "hclust", addrect = 4)
```

## Calculate the correlation matrix for input+output.  
```{r}
cor.matrix <- cor(df)
corrplot::corrplot(cor.matrix, order = "hclust", addrect = 4)
```

## Hierarchical clustering dendogram
```{r}
fviz_dend(link.ward.D2, k = 4, rect = TRUE)
```


## CLUSTERING BASED ON BOTH OUTPUT SUMMARY STATISTICS

## Select optimum number of clusters
```{r}
fviz_nbclust(NbClust(df_output, distance = "euclidean", index = "all", method = "ward.D2"))
```

## Hierarchical clustering
```{r}
d <- dist(df_output, method = "euclidean")
link.ward.D2 <- hclust(d, method = "ward.D2")
```

## Input boxplot by cluster
```{r}
clust.3 <- cutree(link.ward.D2, k = 3)
ggplot(data = reshape2::melt(pareto.IO %>% select("Ax", "Bx", "Dx", "Cx", "Fx", "Ex") %>% mutate(row.ID = 1:nrow(pareto.IO)),
                             id.vars = c("row.ID")) %>% 
           left_join(.,
                     data.frame(row.ID = 1:nrow(df),
                                cluster = clust.3),
                     by = c("row.ID"))) + 
    geom_boxplot( mapping = aes(x = cluster, y=value, fill = as.factor(cluster))) + 
    facet_wrap(~variable, scales = "free") + 
    scale_fill_discrete("Cluster")+theme_bw()
```

## Output boxplot by cluster
```{r}
clust.3 <- cutree(link.ward.D2, k = 3)
ggplot(data = reshape2::melt(pareto.IO %>% select("Gx","Hx", "Ix", "Jx","Kx","Lx", "Mx", "Nx") %>% mutate(row.ID = 1:nrow(pareto.IO)),
                             id.vars = c("row.ID")) %>% 
           left_join(.,
                     data.frame(row.ID = 1:nrow(df),
                                cluster = clust.3),
                     by = c("row.ID"))) + 
    geom_boxplot( mapping = aes(x = cluster, y=value, fill = as.factor(cluster))) + 
    facet_wrap(~variable, scales = "free") + 
    scale_fill_discrete("Cluster")+theme_bw()
```

## Matrix of scatterplots for visual inspection
```{r}
matrix.df <-df_output
matrix.df$cluster <- as.factor(clust.3)
ggpairs(matrix.df,columns = 1:8, bins=10, aes(colour = cluster))
```



## Calculate the correlation matrix for output.  
```{r}
cor.matrix <- cor(df_output)
corrplot::corrplot(cor.matrix, order = "hclust", addrect = 3)
```

## Calculate the correlation matrix for input+output.  
```{r}
cor.matrix <- cor(df)
corrplot::corrplot(cor.matrix, order = "hclust", addrect = 3)
```

## Hierarchical clustering dendogram
```{r}
fviz_dend(link.ward.D2, k = 3, rect = TRUE)
```



## CLUSTERING BASED ON BOTH OUTPUT SUMMARY STATISTICS, WITH K=4

## Hierarchical clustering
```{r}
d <- dist(df_output, method = "euclidean")
link.ward.D2 <- hclust(d, method = "ward.D2")
```

## Input boxplot by cluster
```{r}
clustAll.4 <- cutree(link.ward.D2, k = 4)
ggplot(data = reshape2::melt(pareto.IO %>% select("Ax", "Bx", "Dx", "Cx", "Fx", "Ex") %>% mutate(row.ID = 1:nrow(pareto.IO)),
                             id.vars = c("row.ID")) %>% 
           left_join(.,
                     data.frame(row.ID = 1:nrow(df),
                                cluster = clustAll.4),
                     by = c("row.ID"))) + 
    geom_boxplot( mapping = aes(x = cluster, y=value, fill = as.factor(cluster))) + 
    facet_wrap(~variable, scales = "free") + 
    scale_fill_discrete("Cluster")+theme_bw()
```

## Output boxplot by cluster
```{r}
ggplot(data = reshape2::melt(pareto.IO %>% select("Gx","Hx", "Ix", "Jx","Kx","Lx", "Mx", "Nx") %>% mutate(row.ID = 1:nrow(pareto.IO)),
                             id.vars = c("row.ID")) %>% 
           left_join(.,
                     data.frame(row.ID = 1:nrow(df),
                                cluster = clustAll.4),
                     by = c("row.ID"))) + 
    geom_boxplot( mapping = aes(x = cluster, y=value, fill = as.factor(cluster))) + 
    facet_wrap(~variable, scales = "free") + 
    scale_fill_discrete("Cluster")+theme_bw()
```

## Matrix of scatterplots for visual inspection
```{r}
matrix.df <-df_output
matrix.df$cluster <- as.factor(clustAll.4)
ggpairs(matrix.df,columns = 1:8, bins=10, aes(colour = cluster))
```


## Calculate the correlation matrix for output.  
```{r}
cor.matrix <- cor(df_output)
corrplot::corrplot(cor.matrix, order = "hclust", addrect = 3)
```

## Calculate the correlation matrix for input+output.  
```{r}
cor.matrix <- cor(df)
corrplot::corrplot(cor.matrix, order = "hclust", addrect = 3)
```

## Hierarchical clustering dendogram
```{r}
fviz_dend(link.ward.D2, k = 4, rect = TRUE)
```



## 3D Plots to visualize Pareto Design
```{r}
clustered.df <-pareto.IO
clustered.df$cluster <- as.factor(clust.3)
plot_ly(clustered.df, x = ~Bx, y = ~Ax, z =~Dx, marker = list(color = ~cluster, colorscale = c('#FFE1A1', '#683531'), showscale = TRUE)) %>%
    add_markers() %>%
    layout(scene = list(xaxis = list(title = 'Ax'),
                        yaxis = list(title = 'Bx'),
                        zaxis = list(title = 'Dx')),
            annotations = list(
           x = 1.13,
           y = 1.05,
           text = 'Cluster',
           xref = 'paper',
           yref = 'paper',
           showarrow = FALSE
         ))
```

```{r}
plot_ly(clustered.df, x = ~Cx, y = ~Fx, z =~Ex, marker = list(color = ~cluster, colorscale = c('#FFE1A1', '#683531'), showscale = TRUE)) %>%
    add_markers() %>%
    layout(scene = list(xaxis = list(title = 'Cx'),
                        yaxis = list(title = 'Fx'),
                        zaxis = list(title = 'Ex')),
            annotations = list(
           x = 1.13,
           y = 1.05,
           text = 'Cluster',
           xref = 'paper',
           yref = 'paper',
           showarrow = FALSE
         ))
```

```{r}
plot_ly(clustered.df, x = ~Hx, y = ~Ix, z =~Jx, marker = list(color = ~cluster, colorscale = c('#FFE1A1', '#683531'), showscale = TRUE)) %>%
    add_markers() %>%
    layout(scene = list(xaxis = list(title = 'Hx'),
                        yaxis = list(title = 'Ix'),
                        zaxis = list(title = 'Jx')),
            annotations = list(
           x = 1.13,
           y = 1.05,
           text = 'Cluster',
           xref = 'paper',
           yref = 'paper',
           showarrow = FALSE
         ))
```